---
title: "HW 11"
author: "dani smith"
date: "2023-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figure: Income Category and Health Status
```{r, Message = FALSE, warning= FALSE}
library(tidycensus)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(grDevices)
library(scales)
library(RColorBrewer)

BRFFS_2021 <- read_csv("/Users/iamda/Downloads/LLCP2021.csv")

BRFFS_2021 <- BRFFS_2021|>
  select(
    `_SEX`,`_STATE`,`_AGEG5YR`,`_RACE`,`INCOME3`,`EDUCA`,`GENHLTH`,
    `_HLTHPLN`,`_RFBMI5`,`CVDINFR4`,`CVDSTRK3`,`DIABETE4`,`_RFHYPE6`,
    `_RFDRHV7`,`SMOKE100`,`PHYSHLTH`,`MENTHLTH`,`WEIGHT2`,`PREGNANT`)


BRFFS_2021_Inc__Hlth_Filter <- BRFFS_2021|>
  filter(INCOME3<12,GENHLTH<6)

income_bins <- c(
  .5,1.5,2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5,10.5,11.5)
income_labels <- c(
  "<$10k","$10-15k","$15-20k","$20-25k",
  "$25-35k","$35-50k","$50-75k","$75-100k",
  "$100-150k","$150-200k","$200k+")

BRFFS_2021_Inc__Hlth_Filter$INCOME3 <- as.numeric(BRFFS_2021_Inc__Hlth_Filter$INCOME3)

BRFFS_2021_Inc__Hlth_Filter$income_group <- cut(
  BRFFS_2021_Inc__Hlth_Filter$INCOME3, breaks = income_bins, labels = income_labels, right = FALSE)

total_counts <- BRFFS_2021_Inc__Hlth_Filter|>
  group_by(income_group) |>
  count()


BRFFS_2021_Inc__Hlth_total <- merge(
  BRFFS_2021_Inc__Hlth_Filter, total_counts, by = "income_group", suffixes = c("", "_total"))

  
BRFFS_2021_Inc__Hlth_total<- BRFFS_2021_Inc__Hlth_total|>
  group_by(income_group,GENHLTH)|>
  mutate(hlth_count = n())|>
  ungroup()

topinc_count <- BRFFS_2021_Inc__Hlth_total|>
  filter(INCOME3 == "11")|>
  group_by(INCOME3)|>
  mutate(inc_count = n())|>
  select(inc_count)

exc_hlth_top_inc_count <- BRFFS_2021_Inc__Hlth_total|>
  filter(INCOME3 == "11",GENHLTH == "1")|>
  mutate(exc_inc_count = n())|>
  select(exc_inc_count)

annotation_value <- exc_hlth_top_inc_count$exc_inc_count/topinc_count$inc_count

#print(annotation_value[1]*100)

  BRFFS_2021_Inc__Hlth_total|>
  na.omit()|>
  ggplot(aes(x = income_group, fill = factor(GENHLTH))) +
  geom_bar(position = "fill", stat = "count") +
  labs(title = "Patient Income vs. General Health", x = "Income Level", y = "Reported Health Status")+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),  
        axis.text.y = element_text(angle = 0, hjust = 1.1),
        axis.ticks.x = element_blank()) +
  guides(fill = guide_legend(title = "Health Ratings"),
         shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)),
         ) +
    scale_fill_manual(values = c("#8FB58F","#B8CC9E","#ECE09E","#E3BE9E","#D4A4A4")
                      ,labels =c("Excellent","Very Good","Good","Fair","Poor"))+
  scale_y_continuous(labels = scales::percent_format(scale = 100)) +
  coord_flip() +
  theme(
    legend.position = c(.95, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 5)
    ) +
  theme_bw() +
  theme(legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(5, 5, 5, 5),
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6))+
      annotate(
    geom = "curve", x = 8, y = .6, xend = 10.9, yend = .8, 
    curvature = .3, arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(geom = "text", x = 8, y = .48, label = "(34.10%)", hjust = "left")
  
```
This figure looks at the proportion of reported general health ratings (from poor to excellent health) of patients within different income levels. The x axis indicates the percentage of total patients within an income level category, while the y axis identifies the income level categories for each bar. Within each stacked bar, you can identify the proportion of different health ratings that patients in that category reported.

Looking at the plot, there's a consistent decrease in the proportion of patients reporting poor, fair, and good health as income level rises. Moreover, there's a noticeable increase in proportions for very good and excellent health as income level rises. This highlights a connection between patients income status and self-reported general health outcomes, with higher incomes potentially impacting better-graded self-reported health outcomes

# Table: Health Status, Insured Status, and Overweight/Obese Outcomes
```{r, message = FALSE,echo = TRUE, warning=FALSE}
library(knitr)
library(kableExtra)
hlth_status_mapping <- c("Excellent","Very Good","Good","Fair","Poor")
hlth_pln_mapping <- c("Insured","UnInsured")
obesity_mapping <- c("No", "Yes")

brffs_obesity_hlth_ins <- BRFFS_2021|>
  select(`_HLTHPLN`,`GENHLTH`,`_RFBMI5`)|>
  filter(`_HLTHPLN` %in% c(1,2), `GENHLTH` %in% c(1,2,3,4,5),`_RFBMI5` %in% c(1,2))

obesity_hlth_ins_rename <- brffs_obesity_hlth_ins|>
  mutate(Health_Status = factor(`GENHLTH`, levels = 1:5, labels = hlth_status_mapping),
         Health_Plan_Enrollment = factor(`_HLTHPLN`, levels = 1:2, labels = hlth_pln_mapping),
         Obesity = factor(`_RFBMI5`, levels = 1:2, labels = obesity_mapping))|>
  select(Health_Status,Health_Plan_Enrollment,Obesity)


format_obesity_hlth_ins <- obesity_hlth_ins_rename|>
  group_by(Health_Status,`Health_Plan_Enrollment`,`Obesity`)|>
  summarise(Count = n())

pct_obesity_health_ins <- format_obesity_hlth_ins|>
  pivot_wider(names_from = Obesity, values_from = Count, values_fill = 0)

obesity_pct_table <-pct_obesity_health_ins|>
  mutate(
    `no_pct` = paste(round(No/(No + Yes) *100,2),"%", sep = " "),
    `yes_pct` = paste(round(Yes/(No + Yes)*100,2),"%", sep = " ")
  )|>
  rename(`Overweight or Obese (No)` = no_pct, `Overweight or Obese (Yes)` = yes_pct)|>
  select(-No,-Yes)
 
 kable(obesity_pct_table, booktabs = TRUE, col.names = c(
    'Health Status', 'Health Enrollment', 'Overweight or Obese (No)', 'Overweight or Obese (Yes)'))|>
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))|>
  row_spec(0,bold = T)|>
  row_spec(c(2,4,6,8),hline_after = T)

```

The table reports the percentages of patient's surveyed as overweight or obese and those surveyed as not overweight or obese, by health status and insured/uninsured status.This is meant to look at the connection between self-reported general health,health insurance enrollment, and overweight/obese outcomes among patients. 

Notably, in all categories of health status and insured status, majority of patients were surveyed as overweight or obese; which appears consistent with American overweight/obesity data. In the table, it's clear that there are lower proportions of patients surveyed as overweight/obese with improvements in health status rating. Additionally, once health status improves from good to very good, the disparity between patients reported as overweight/obese within insured and uninsured groups begins to noticeably decrease. Interestingly, in every health status category, other than 'excellent', uninsured patients have lower rates of patient's reported as overweight/obese.